Understanding Vulnerability: An Analysis of Toronto Police Service's Annual Statistical Report on Crimes Against the Person(2014-2022)

### 2024-01-25 output/paper/paper.qmd
updated paper.qmd and rendered pdf

The dataset for this study was obtained from OpenDataToronto [@opendatatoronto], specifically the "Police Annual Statistical Report - Victims of Crimes" published by the Toronto Police Services. The dataset includes all identified victims of crimes against the person reported to the Toronto Police and the type of crime that was committed along with the sex and age group information of the victim. Here, "crimes against the person" is referring to CCJS level 1 only @victimsofcrimes. As per statistics Canada, examples of CCJS level 1 crimes are assault, sexual assault etc. @statscanada. Our research offers an examination of the trends in the number of reported victims of crimes each year over the span of nine years from 2014-2022, the type of crimes committed (assault, sexual assault, robbery or other) against them and factors that make them vulnerable to these crimes. The examination was executed using the R programming language [@citeR] and enhanced with diverse analytical tools, including Tidyverse [@tidyverse], Janitor [@janitor], and Knitr [@knitr].

@@

## Limitations

-   Age calculated from date of occurrence

-   Filtered by reported year

-   Victims only for CCJS Level 1 = 'Crimes Against the Person' and where the victim name is not null

## **Data Quality**

As of 23 January, 2024, the dataset was last refreshed on 14 January, 2024 with an overall data quality score of 99% and with a gold rating

The dataset has a score of 100% for freshness which means the data set is up-to-date and refreshed on schedule.

The dataset has a score of 100% for metadata and completeness which means the data is well described and complete. The dataset has a score of 93% for Usability due to one column "Category" having constant values.

@@

```{r}
#| eval: true
#| warning: false
#| label: fig-1
#| fig-cap: "Total count of victims of crimes against the person per year (2014-2022)"

@@

```{r}
#| eval: true
#| warning: false
#| label: fig-2
#| fig-cap: "Total count of victims of crimes against the person per year by type of crime (2014-2022)"

@@

```{r}
#| eval: true
#| warning: false
#| label: fig-3
#| fig-cap: "Total count of victims of crimes against the person of each sex per year"

@@

```{r}
#| eval: true
#| warning: false
#| label: fig-4
#| fig-cap: "Total count of victims of crimes against the person of each age group per year"

@@

>>>>

The dataset for this study was obtained from OpenDataToronto [@opendatatoronto], specifically the "Police Annual Statistical Report - Victims of Crimes" published by the Toronto Police Services. The dataset includes all identified victims of crimes against the person reported to the Toronto Police and the type of crime that was committed along with the sex and age group information of the victim. Here, "crimes against the person" is referring to CCJS level 1 only [@victimsofcrimes]. As per statistics Canada, examples of CCJS level 1 crimes are assault, sexual assault etc. @statscanada. Our research offers an examination of the trends in the number of reported victims of crimes each year over the span of nine years from 2014-2022, the type of crimes committed (assault, sexual assault, robbery or other) against them and factors that make them vulnerable to these crimes. The examination was executed using the R programming language [@citeR] and enhanced with diverse analytical tools, including Tidyverse [@tidyverse], Janitor [@janitor], and Knitr [@knitr].

@@

As mentioned above, the dataset utilized is this paper is " Police Annual Statistical Report - Victims of Crimes" published by the Toronto Police Services. The dataset includes all identified victims of crimes against the person, including, but not limited to, those that may have been deemed unfounded after investigation, those that may have occurred outside the City of Toronto limits, or have no verified location. The dataset includes data from 2014-2022 with 1110 samples each with 9 variables. Some samples in variables such as "Sex" and "Age Group" have unknown values. But the number of such samples is relatively and consistently small for each year and hence doesn't create a huge deviation in the data. Another potential concern is the number of unidentified or unreported victims of these crimes. However, since the data is only considering the CCSJ level 1 crimes, it limits the number of unreported CCSJ level 1 crimes to a small number. As per statistics Canada, examples of CCSJ level 1 crimes are assault, sexual assault etc. @statscanada. Moreover, crimes like assault, sexual assault, and robbery usually do not go unreported. Finally, the variable "Subtype" has "other" as one of the options which does not provide enough insight into what the other subtypes of crime are apart from assault, sexual assault, and robbery, thereby limiting the connection that can be drawn between different variables. No other datasets were considered because this data is directly published by Toronto Police Services, the original owner of the data. Additionally, the dataset was rated highly in data quality as below [@victimsofcrimes]:

-   As of 23 January, 2024, the dataset was last refreshed on 14 January, 2024 with an overall data quality score of 99% and with a gold rating.

-   The dataset has a score of 100% for freshness which means the data set is up-to-date and refreshed on schedule.

-    The dataset has a score of 100% for metadata and completeness which means the data is well described and complete.

-   The dataset has a score of 93% for Usability due to one column "Category" having constant values.

In order to make the dataset manageable, unnecessary variables such as ID (served no purpose in our analysis) and Category (all samples had the same value, "Crimes Against the Person) were excluded.

## 2.2 Variable of Interest

The data includes the year the victim was reported, the subtype of the crime committed against the victim which was broken into four sub-categories "assault", "sexual assault", "robbery" and "other".

The victim is then identified as "Male", "Female" or "Unknown" under the sex variable. This is followed by variable age group and age cohort which are connected in the sense that ages "\<12" relates to age group "Child", ages "12-17" relates to age group "Youth" and all other ages relates to age group "Adult". Apart from these, if age group is "Unknown", age cohort is "Unknown" as well and vice versa. Finally, there is a variable to count the number of victims reported per year per subtype per sex per age group per age cohort which together makes one sample on the dataset. This variable is an important aspect of measurement as it represents the number of occurrences for each combination of all other variables hence allowing a quantitative analysis of those variables.

@@

```{r}
#| eval: true
#| warning: false
#| label: fig-1
#| fig-cap: "Graph of total number of victims per year (2014-2022)"

@@

```{r}
#| eval: true
#| warning: false
#| label: fig-2
#| fig-cap: "Graph of total number of victims per year by type of crime (2014-2022)"

@@

```{r}
#| eval: true
#| warning: false
#| label: fig-3
#| fig-cap: "Graph of total number of victims per year by sex"

@@

```{r}
#| eval: true
#| warning: false
#| label: fig-4
#| fig-cap: "Graph of total number of victims per year by age group"

@@

```{r}
#| eval: true
#| warning: false
#| label: fig-5
#| fig-cap: "Graph of total number of adult victims per year by age cohort"

setwd("/cloud/project")

# Read data
crime_data <- read_csv("output/data/edited_data.csv", show_col_types = FALSE) 
column_specifications <- spec(crime_data)

# Define the order of age cohorts
age_order <- c("18 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65+")

# Filter out unwanted age cohorts
filtered_data <- subset(crime_data, !(AGE_COHORT %in% c("<12", "Unknown", "12 to 17")))

# Aggregate total counts by REPORT_YEAR, AGE_COHORT, and sum
age_counts <- aggregate(COUNT_ ~ REPORT_YEAR + AGE_COHORT, data = filtered_data, sum)

# Create a line graph using ggplot2 with age cohort on the x-axis
ggplot(age_counts, aes(x = as.factor(REPORT_YEAR), y = COUNT_, color = AGE_COHORT, group = AGE_COHORT)) +
  geom_line(linewidth = 1, alpha = 0.7) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Year",
    y = "Total Number of Crimes"
  ) +
  scale_color_manual(values = c(
    "18 to 24" = "blue", "25 to 34" = "red", "35 to 44" = "green", 
    "45 to 54" = "purple", "55 to 64" = "orange", "65+" = "pink"
  )) +  # Specify colors
  theme_minimal()
